[Re] Non-additive coupling enables propagation of synchronous spiking activity in purely random networks

### 2018-03-08 article/Caze_2018.tex
Update Caze_2018.tex
Need to compile the tex file I changed the thing in four locations (as we saw). Also need to update the rebuttal.doc to refer to the changes (line ref in tex)

@@

\subsection*{Semi-analytical and analytical methods}
We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}).
To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows. First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.  We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential. From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
\begin{equation}
F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
\end{equation}
Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e. $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$. Let us assume that in the last iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e. that $N-g_i$ neurons are available to spike in this iteration. To calculate the probability for a single (non-refractory) neuron $j$ to spike in response to these $g_i$ neurons, we will first calculate the probability that out of the $g_i$ neurons that spiked, exactly $n_\text{ex}$ neurons are connected with excitatory connections and $n_\text{in}$ neurons are connected with inhibitory connections to neuron $j$. There are $\binom{g_i}{n_\text{ex} + n_\text{in}}$ possibilities to chose $n_\text{ex} + n_\text{in}$ out of $g_i$ neurons, and $\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}}$ possibilities to divide these neurons into groups of size $n_\text{ex}$ and $n_\text{in}$. This yields the total number of combinations $c(g_i, n_\text{ex}, n_\text{in})$:
\begin{equation}
c(g_i, n_\text{ex}, n_\text{in}) = \binom{g_i}{n_\text{ex} + n_\text{in}}\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}} = \frac{g_i!}{n_\text{ex}!n_\text{in}!(g_i - n_\text{ex} - n_\text{in})!}
\end{equation}
The probability that these selected $n_\text{ex}$ neurons are actually connected via excitatory connections to neuron $j$, that the $n_\text{in}$ neurons are actually connected via inhibitory connections to neuron $j$, and that the remaining $g_i - n_\text{ex} - n_\text{in}$ neurons are \emph{not} connected to neuron $j$ is:
\begin{equation}
p(g_i, n_\text{ex}, n_\text{in}) = (p_0 p_\text{ex})^{n_\text{ex}} (p_0 p_\text{in})^{n_\text{in}} (1 - p_0)^{g_i - n_\text{ex} - n_\text{in}}
\end{equation}
Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike. This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
\begin{equation}
P^s(g_i) = \sum_{n_\text{ex}=1}^{g_i}\sum_{n_\text{in}=0}^{g_i-n_\text{ex}} F\big(\epsilon(n_\text{ex}, n_\text{in})\big) c(g_i, n_\text{ex}, n_\text{in}) p(g_i, n_\text{ex}, n_\text{in})
\end{equation}
In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, i.e.:
\begin{equation}
E(g_{i+1}|g_i = g_i') = (N - g_i') P^s(g_i')
\end{equation}

@@

@@

We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences. These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision. We tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained qualitatively the same results (not shown). Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.

>>>>

Our replication present slight differences with the original grid search. The stable region surface ends up earlier in a regime with strong inhibitory connections. The odd blue-yellow-red region presents in the original article differs from the continuous transition we obtained in our simulations. Implementation details might explain these differences

@@

\subsection*{Semi-analytical and analytical methods}
We did not attempt to replicate the analytical solution for the evolution of synchronous pulses based on diffusion approximation (blue dots in Fig.~4 of \cite{Memmesheimer2012}). The article explains in more details the semi-analytical work that was more straightforward to reproduce.
To calculate the semi-analytical solutions for Fig.~\ref{fig:markov}, based on a numerical estimate of the distribution $P(V)$, we proceeded as follows. First, we ran $2\times 50$ (once for the linear, once for the non-linear network) simulations of \SI{250}{\milli\second} each of a network with the default parameters but without any additional synchronous stimulation.  We binned all membrane potential values over all time steps and neurons into 100 bins in the voltage interval $[-\Theta/8, \Theta]$, to obtain an estimate of $P(V)$, the distribution of the membrane potential. From this distribution we can estimate the probability $F$ that an additional input $\epsilon$ drives the neuron over its threshold (or equivalently, that the neuron is less than $\epsilon$ away from its threshold):
\begin{equation}
F(\epsilon) = P(V>\Theta-\epsilon) = \int_{\Theta-\epsilon}^{\Theta}P(V)dV
\end{equation}
Let $\epsilon(n_\text{ex}, n_\text{in})$ be the total input to a neuron that receives $n_\text{ex}$ excitatory and $n_\text{in}$ inhibitory spikes, i.e. $\epsilon(n_\text{ex}, n_\text{in}) = \sigma(n_\text{ex}w_\text{ex}) + n_\text{in}w_\text{in}$. Let us assume that in the last iteration $g_i$ neurons spiked and that these neurons are refractory now, i.e. that $N-g_i$ neurons are available to spike in this iteration. To calculate the probability for a single (non-refractory) neuron $j$ to spike in response to these $g_i$ neurons, we will first calculate the probability that out of the $g_i$ neurons that spiked, exactly $n_\text{ex}$ neurons are connected with excitatory connections and $n_\text{in}$ neurons are connected with inhibitory connections to neuron $j$. There are $\binom{g_i}{n_\text{ex} + n_\text{in}}$ possibilities to chose $n_\text{ex} + n_\text{in}$ out of $g_i$ neurons, and $\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}}$ possibilities to divide these neurons into groups of size $n_\text{ex}$ and $n_\text{in}$. This yields the total number of combinations $c(g_i, n_\text{ex}, n_\text{in})$:
\begin{equation}
c(g_i, n_\text{ex}, n_\text{in}) = \binom{g_i}{n_\text{ex} + n_\text{in}}\binom{n_\text{ex} + n_\text{in}}{n_\text{ex}} = \frac{g_i!}{n_\text{ex}!n_\text{in}!(g_i - n_\text{ex} - n_\text{in})!}
\end{equation}
The probability that these selected $n_\text{ex}$ neurons are actually connected via excitatory connections to neuron $j$, that the $n_\text{in}$ neurons are actually connected via inhibitory connections to neuron $j$, and that the remaining $g_i - n_\text{ex} - n_\text{in}$ neurons are \emph{not} connected to neuron $j$ is:
\begin{equation}
p(g_i, n_\text{ex}, n_\text{in}) = (p_0 p_\text{ex})^{n_\text{ex}} (p_0 p_\text{in})^{n_\text{in}} (1 - p_0)^{g_i - n_\text{ex} - n_\text{in}}
\end{equation}
Finally, we need to sum these probabilities up over all possible values for $n_\text{ex}$ and $n_\text{in}$, multiplying it with the probability that the generated input elicits a spike. This yiels the total probability $P^s(g_i)$ that a neuron $j$ spikes in response to the activity of $g_i$ neurons in the previous iteration:
\begin{equation}
P^s(g_i) = \sum_{n_\text{ex}=1}^{g_i}\sum_{n_\text{in}=0}^{g_i-n_\text{ex}} F\big(\epsilon(n_\text{ex}, n_\text{in})\big) c(g_i, n_\text{ex}, n_\text{in}) p(g_i, n_\text{ex}, n_\text{in})
\end{equation}
In Fig.~\ref{fig:markov}, we plot the expected value of spiking neurons during the next iteration, i.e.:
\begin{equation}
E(g_{i+1}|g_i = g_i') = (N - g_i') P^s(g_i')
\end{equation}

@@

The result of our simulation seems shifted upward compared to the original work. This difference is most visible at the origin of the plot. This might come from "false positive" spikes generated by the background activity (~5 spikes more in our simulation results). Yet our work strengthen the original result demonstrating a qualitative difference between linearly coupled and non-linearly coupled networks.

@@

We managed to qualitatively reproduce the main results of Raoul-Martin Memmes\-heimer and Marc Timme \cite{Memmesheimer2012} despite some slight quantitative differences. The grid search differs in certain regions and the markov analysis demonstrate a slight shifth upward for the numerical results. These difference could be due to our use of clock-based instead of event-driven simulations, as it forces all activity to a temporal grid of finite precision. We tried finer (\SI{0.05}{\milli\second}) and coarser (\SI{0.2}{\milli\second}, \SI{0.5}{\milli\second}) temporal resolutions and obtained qualitatively the same results (not shown). Another potential source for the differences is that our initial conditions do not include spikes ``in transit'' at the start of the simulation, which might further contribute to the quantitative differences we see.
